---
title: "MIS-64040-001"
author: "Kiran Kour"
date: "2022-09-08"
output: html_document
---

# Assignment 1 : Know how to use software tools (such as R) effectively to implement machine learning algorithms for data mining/visualization and analytics

#The source of my dataset is (<https://www.contextures.com/xlsampledata01.html>) which you can access from the above link.

**#Importing the dataset:**

```{r}
foodsales <- read.csv("foodsales.csv")
head(foodsales)
```

**#Descriptive Statistics for Categorical Variables:**

```{r}
freq <- table(foodsales$City)
head(freq)
```

```{r}
freq <- table(foodsales$Product,foodsales$Category)
head(freq)
```

```{r}
freq <- table(foodsales$OrderDate,foodsales$Category,foodsales$Product)
head(freq)
```

**#Plotting of Qualitative Variables:**

```{r}
freq <- table(foodsales$City,foodsales$Product)
barplot(freq,main="Country VS Product",xlab="Country",ylab="Product",horiz=TRUE,col="blue",las=1,cex.names=0.5)
```

**#Descriptive Statistics for Quantitative Variables:**

```{r}
summary(foodsales)
```

**#Summary measures for Total Price:**

```{r}
mean(foodsales$TotalPrice)
median(foodsales$TotalPrice)
sd(foodsales$TotalPrice)
range(foodsales$TotalPrice)
```

```{r}
require(e1071)
skewness(foodsales$TotalPrice)
kurtosis(foodsales$TotalPrice)
```

**#Plotting of Quantitative Variables:**

```{r}
plot(foodsales$Quantity~foodsales$UnitPrice,main="Scatter Plot",xlab="UnitPrice",ylab="Quantity",col="Black",type="p")
```

```{r}
plot(foodsales$Quantity~foodsales$TotalPrice,main="Scatter Plot",xlab="TotalPrice",ylab="Quantity",pch=19,frame=FALSE)
abline(lm(Quantity~TotalPrice,data=foodsales),col="blue")
```

**#Transformation of the Variable:**

```{r}
sqrt_UnitPrice<- sqrt(foodsales$UnitPrice)
hist(foodsales$UnitPrice,col="Steel blue",main="Original")
```

**#Plot for Square Root Transformation:**

```{r}
hist(sqrt_UnitPrice,col="coral2",main="Square Root Transformed")
```
