---
title: "MIS-64060-001(A3)"
author: "Kiran Kour"
date: "2022-10-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing required packages

```{r}
#install.packages("reshape")
#install.packages("reshape2")
#install.packages("melt")
#install.packages("naivebayes")
#install.packages("pROC")
```
```{r}
library(readr)
library(dplyr)
library(caret)
library(class)
library(melt)
library(reshape)
library(reshape2)
library(ggplot2)
library(ISLR)
library(naivebayes)
library(e1071)
library(pROC)
```

#Importing the dataset

```{r}
universalbank<- read.csv("UniversalBank.csv")
head(universalbank,n=5)
```
#Checking for missing values using is.na()

```{r}
bank <- is.na.data.frame("universalbank")
```

#Converting the data type of categorical variables to factor

```{r}
universalbank$Personal.Loan= as.factor(universalbank$Personal.Loan)
universalbank$Online= as.factor(universalbank$Online)
universalbank$CreditCard= as.factor(universalbank$CreditCard)
```

#Data Partition and Normalization

```{r}
set.seed(333)
Train_Index<- createDataPartition(universalbank$Personal.Loan, p=0.6, list=FALSE)
Train <-universalbank[Train_Index,]
Valid <-universalbank[-Train_Index,]
```

```{r}
Model_norm <- preProcess(Train[,-c(10,13:14)],method = c("center", "scale"))
Train_norm <- predict(Model_norm,Train)
Valid_norm<- predict(Model_norm,Valid)
```

# Part A: Creating pivot table for Training data

```{r}
Table.OCP <- table(Train_norm$Personal.Loan, Train_norm$Online, Train_norm$CreditCard, dnn=c("Personal Loan", "Online", "Credit Card"))
Table.OCP
```

## Part B: Computing P(Loan | Online & CC)

As we look that the pivot table created in part A out of the total 550 records where of active online banking users with credit cards, 49 had accepted a personal loan, so $$\textbf{P}(\text{Loan}=1\ |\ \text{CC}=1\text {and}\text{ Online}=1) = \frac{49}{550} = 0.089$$.
```{r}
# Computing P(loan | Online & CC)
Table.OCP[2,2,2] / (Table.OCP[2,2,2] + Table.OCP[1,2,2])
```

# Part C: Creating two separate pivot tables for Training data.One will have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC. 

```{r}
Table_Online <- table(Train_norm$Personal.Loan, Train_norm$Online, dnn=c("Personal Loan", "Online"))
Table_Online

Table_CreditCard <- table(Train_norm$Personal.Loan, Train$CreditCard, dnn=c("Personal Loan", "Credit Card"))
Table_CreditCard
```

# Part D : Computing the following quantities:

i.) $$\textbf{P}\text{(CC = 1 | Loan = 1)}\text{= 80/80+208}$$

```{r}
prob_CCL <- Table_CreditCard[2,2] / (Table_CreditCard[2,2] + Table_CreditCard[2,1])
prob_CCL
```

ii.)$$\textbf{P}\text{(Online = 1 | Loan = 1)}\text{= 184/184+104}$$

```{r}
prob_OL <- Table_Online[2,2] / (Table_Online[2,2] + Table_Online[2,1])
prob_OL
```

iii.) $$\textbf{P}\text{(Loan = 1)}\text{= 288/288+2712}$$

```{r}
prob_Loan <- sum(Train_norm$Personal.Loan==1) / length(Train_norm$Personal.Loan)
prob_Loan
```

iv.) $$\textbf{P}\text{(CC = 1 | Loan = 0)}\text{= 806/806+1906}$$

```{r}
prob_CCNL <-Table_CreditCard[1,2] / (Table_CreditCard[1,2] + Table_CreditCard[1,1])
prob_CCNL
```
v.) $$\textbf{P}\text{(Online = 1 | Loan = 0)}\text{= 1621/1621+1091}$$

```{r}
prob_ONL <- Table_Online[1,2] / (Table_Online[1,2] + Table_Online[1,1])
prob_ONL
```
vi.) $$\textbf{P}\text{(Loan = 0)}\text{= 2712/2712+288}$$

```{r}
prob_NL <- sum(Train_norm$Personal.Loan==0) / length(Train_norm$Personal.Loan)
prob_NL
```

# Part E : Using the quantities computed above to compute the Naive Bayes probability P(Loan = 1 | CC = 1, Online = 1).
$$\textbf{P}(\text{Loan}=1\ |\ \text{CC}=1,\text{ Online}=1)\text{= (0.6388 x 0.2777 x 0.096) / (0.6388 x 0.2777 x 0.096 + 0.5977 x 0.2972 x 0.904)= 0.0959}$$

```{r}
(prob_OL * prob_CCL * prob_Loan) / (prob_OL * prob_CCL * prob_Loan + prob_ONL * prob_CCNL * prob_NL)
```

# Part F : Comparing the value obtained from part Naive bayes probability with the one obtained from the pivot table in (B). 

Using the Naive Bayes classifier, we get a higher value for $\text{P}(\text{Loan}=1\ |\ \text{CC}=1,\text{ Online}=1)$ than with the direct computation obtained in part B. Interestingly, in part D we got the value of  $\textbf{P}\text{(Loan = 1)}$ as 0.096 and also in Naive bayes classifer we got the value as 0.096. So the Naive Bayes approach suggests that the probability a person will accept the loan is independent of whether that person is an online user with a bank-issued credit card.

# Part G : Running Naive Bayes on the data

```{r}
naive <- naiveBayes(Personal.Loan~Online+CreditCard,data=Train_norm)
naive
```
The value that is obtained for the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1) from running Naive bayes is 0.09591693 , which is equal to the value derived from part E.

# AUC Value and ROC Curve

```{r}
Predicted_labels <-predict(naive,Valid_norm, type = "raw")
head(Predicted_labels)
```
```{r}
roc(Valid_norm$Online, Predicted_labels[,2])
```

```{r}
plot.roc(Valid_norm$Online,Predicted_labels[,2])
```